---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

# mlr3proba

Probabilistic Supervised Learning for **[mlr3](https://github.com/mlr-org/mlr3/)** ([website](https://mlr3proba.mlr-org.com/)).

<!-- badges: start -->
[![R-CMD-check](https://github.com/mlr-org/mlr3proba/actions/workflows/r-cmd-check.yml/badge.svg)](https://github.com/mlr-org/mlr3proba/actions/workflows/r-cmd-check.yml)
[![runiverse](https://mlr-org.r-universe.dev/badges/mlr3proba)](https://mlr-org.r-universe.dev/mlr3proba)
[![GitHub Discussions](https://img.shields.io/github/discussions/mlr-org/mlr3proba?logo=github&label=Discussions%20Q%26A&color=FFE600)](https://github.com/mlr-org/mlr3proba/discussions)
[![Article](https://img.shields.io/badge/Article-10.1093%2Fbioinformatics%2Fbtab039-brightgreen)](https://doi.org/10.1093/bioinformatics/btab039)
[![StackOverflow](https://img.shields.io/badge/stackoverflow-mlr3-orange.svg?color=pink)](https://stackoverflow.com/questions/tagged/mlr3)
[![Mattermost](https://img.shields.io/badge/chat-mattermost-orange.svg?color=pink)](https://lmmisld-lmu-stats-slds.srv.mwn.de/mlr_invite/)
<!-- badges: end -->

## What is mlr3proba?

`mlr3proba`  is a machine learning toolkit for **probabilistic supervised learning** within the **[mlr3](https://github.com/mlr-org/mlr3)** ecosystem.

It currently supports:

1. **Predictive survival analysis** for single-event tasks with right-censoring.
2. **Unsupervised density estimation** for single variable tasks.
3. **Probabilistic supervised regression**: Supervised regression with a predictive distribution as the return type.

## Feature Overview

Key features of `mlr3proba` focus on survival analysis and are:

- Task frameworks for survival analysis (`TaskSurv`)
- A comprehensive **selection of survival learners** (mostly via
  [mlr3extralearners](https://github.com/mlr-org/mlr3extralearners/))
- A unified `$train()`/`$predict()` model interface to any probabilistic predictive model (frequentist, Bayesian, Deep Learning, or other)
- Use of the **[distr6](https://github.com/alan-turing-institute/distr6)** interface for the survival probability distribution prediction
- A comprehensive selection of **measures** for evaluating the performance of survival learners, with respect to prognostic index (continuous rank) prediction, and probabilistic (distribution) prediction
- Basic **ML pipeline building** integrated with **[mlr3pipelines](https://github.com/mlr-org/mlr3pipelines)** (e.g. transform prediction types)
- **Reduction strategies** to transform survival to classification/regression problems

## Installation

`mlr3proba` is not currently on CRAN.
Please follow one of the two following methods to install it:

### R-universe

Install the latest released version:

```{r eval = FALSE}
install.packages("mlr3proba", repos = "https://mlr-org.r-universe.dev")
```

### GitHub

Install the latest development version:

```{r eval = FALSE}
# install.packages("pak")
pak::pak("mlr-org/mlr3proba")
```

## Learners

- [Core learners](https://mlr3proba.mlr-org.com/reference/index.html#survival-learners) are implemented in `mlr3proba` and include the Kaplan-Meier Estimator, the Cox Proportional Hazards model and the Survival Tree learner.
- In [mlr3extralearners](https://github.com/mlr-org/mlr3extralearners) we have interfaced several more advanced ML learners.
Use the [interactive search table](https://mlr-org.com/learners.html) to search for the available survival learners and see the [learner status page](https://mlr3extralearners.mlr-org.com/articles/learner_status.html) for their live status.

## Measures

For density estimation and probabilistic regression only the **log-loss** is currently implemented.
For survival analysis, see list [here](https://mlr3proba.mlr-org.com/reference/index.html#survival-measures).

Some commonly used measures for right-censored single-event tasks are the following:

| ID | Measure | Package | Category | Prediction Type
| :--| :------ | :------ | :------ | :------- |
| [surv.dcalib](https://mlr3proba.mlr-org.com/reference/mlr_measures_surv.dcalib.html) | D-Calibration | `mlr3proba` | Calibration | `distr`
| [surv.calib_index](https://mlr3proba.mlr-org.com/reference/mlr_measures_surv.calib_index.html) | One-point Calibration | `mlr3proba` | Calibration | `distr`
| [surv.cindex](https://mlr3proba.mlr-org.com/reference/mlr_measures_surv.cindex.html) | Concordance Index | `mlr3proba` | Discrimination | `crank`
| [surv.uno_auc](https://mlr3proba.mlr-org.com/reference/mlr_measures_surv.uno_auc.html) | Uno's AUC | `survAUC` | Discrimination | `lp`
| [surv.graf](https://mlr3proba.mlr-org.com/reference/mlr_measures_surv.graf.html) | Integrated Brier Score | `mlr3proba` | Scoring Rule | `distr`
| [surv.rcll](https://mlr3proba.mlr-org.com/reference/mlr_measures_surv.rcll.html) | Right-Censored Log loss | `mlr3proba` | Scoring Rule | `distr`
| [surv.intlogloss](https://mlr3proba.mlr-org.com/reference/mlr_measures_surv.intlogloss.html) | Integrated Log-Likelihood | `mlr3proba` | Scoring Rule | `distr`

## Bugs, Questions, Feedback

**mlr3proba** is a free and open source software project that encourages participation and feedback.
If you have any issues, questions, suggestions or feedback, please do not hesitate to open an
"issue" about it on the [GitHub page](https://github.com/mlr-org/mlr3proba/issues)\!

In case of problems / bugs, it is often helpful if you provide a "minimum working example" using [reprex](https://reprex.tidyverse.org/) that showcases the behavior.

## Similar Projects

Predecessors to this package are previous instances of survival modelling in **[mlr](https://github.com/mlr-org/mlr)**.
The **[skpro](https://github.com/alan-turing-institute/skpro)** package in the python/scikit-learn ecosystem follows a similar interface for probabilistic supervised learning and is an architectural predecessor.
Several packages exist which allow probabilistic predictive modelling with a Bayesian model specific general interface, such as **[rjags](https://CRAN.R-project.org/package=rjags)** and **[stan](https://github.com/stan-dev/rstan)**.
For implementation of a few survival models and measures, a central package is **[survival](https://github.com/therneau/survival)**.
There does not appear to be a package that provides an architectural framework for distribution/density estimation, see **[this list](https://vita.had.co.nz/papers/density-estimation.pdf)** for a review of density estimation packages in `R`.

## Acknowledgements

Several people contributed to the building of `mlr3proba`.
Firstly, thanks to Michel Lang for writing `mlr3survival`.
Several learners and measures implemented in `mlr3proba`, as well as the prediction, task, and measure surv objects, were written initially in `mlr3survival` before being absorbed into `mlr3proba`.
Secondly thanks to Franz Kiraly for major contributions towards the design of the proba-specific parts of the package, including compositors and predict types.
Also for mathematical contributions towards the scoring rules implemented in the package.
Finally thanks to Bernd Bischl and the rest of the mlr core team for building `mlr3` and for many conversations about the design of `mlr3proba`.

## Citing mlr3proba

If you use mlr3proba, please cite our [Bioinformatics article](https://doi.org/10.1093/bioinformatics/btab039):
```{r echo = FALSE, comment = ""}
toBibtex(citation("mlr3proba"))
```
