---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

# mlr3proba

Package website: [release](https://mlr3proba.mlr-org.com/)

Probabilistic Supervised Learning for **[mlr3](https://github.com/mlr-org/mlr3/)**.

<!-- badges: start -->
[![r-cmd-check](https://github.com/mlr-org/mlr3proba/actions/workflows/r-cmd-check.yml/badge.svg)](https://github.com/mlr-org/mlr3proba/actions/workflows/r-cmd-check.yml)
[![cran checks](https://cranchecks.info/badges/worst/mlr3proba)](https://cran.r-project.org/web/checks/check_results_mlr3proba.html)
[![Article](https://img.shields.io/badge/Article-10.1093%2Fbioinformatics%2Fbtab039-brightgreen)](https://doi.org/10.1093/bioinformatics/btab039)
[![StackOverflow](https://img.shields.io/badge/stackoverflow-mlr3-orange.svg)](https://stackoverflow.com/questions/tagged/mlr3)
[![Mattermost](https://img.shields.io/badge/chat-mattermost-orange.svg)](https://lmmisld-lmu-stats-slds.srv.mwn.de/mlr_invite/)
<!-- badges: end -->

## What is mlr3proba ?

**mlr3proba** is a machine learning toolkit for making probabilistic predictions within the **[mlr3](https://github.com/mlr-org/mlr3)** ecosystem. It currently supports the following tasks:

* **Probabilistic supervised regression** - Supervised regression with a predictive distribution as the return type.
* **Predictive survival analysis** - Survival analysis where individual predictive hazards can be queried. This is equivalent to probabilistic supervised regression with censored observations.
* **Unconditional distribution estimation**, where the distribution is returned. Sub-cases are density estimation and unconditional survival estimation.

Key features of **mlr3proba** are

* A unified fit/predict model interface to any probabilistic predictive model (frequentist, Bayesian, or other)
* Pipeline/model composition
* Task reduction strategies
* Domain-agnostic evaluation workflows using task specific algorithmic performance measures.

**mlr3proba** makes use of the **[distr6](https://github.com/alan-turing-institute/distr6)** probability distribution interface as its probabilistic predictive return type.


## Feature Overview

The current **mlr3proba** release focuses on survival analysis, and contains:

* Task frameworks for survival analysis (`TaskSurv`)
* A comprehensive selection of predictive survival learners (mostly via [mlr3extralearners](https://github.com/mlr-org/mlr3extralearners/))
* A comprehensive selection of performance measures for predictive survival learners, with respect to prognostic index (continuous rank) prediction, and probabilistic (distribution) prediction
* PipeOps integrated with **[mlr3pipelines](https://github.com/mlr-org/mlr3pipelines)**, for basic pipeline building, and reduction/composition strategies using linear predictors and baseline hazards.

## Roadmap

The vision of **mlr3proba** is to provide comprehensive machine learning functionality to the mlr3 ecosystem for continuous probabilistic return types.

The lifecycle of the survival task and features are considered `maturing` and any major changes are unlikely.

The density and probabilistic supervised regression tasks are currently in the early stages of development. Task frameworks have been drawn up, but may not be stable; learners need to be interfaced, and contributions are very welcome (see [issues](https://github.com/mlr-org/mlr3proba/issues)).


## Installation

`mlr3proba` is not on CRAN and is unlikely to be reuploaded (see [here](https://twitter.com/RaphaelS101/status/1506321623250571265) for reasons). As such you must install with one of the following methods:

### Install from r-universe:

```{r eval=FALSE}
options(repos=c(
  mlrorg = 'https://mlr-org.r-universe.dev',
  raphaels1 = 'https://raphaels1.r-universe.dev',
  CRAN = 'https://cloud.r-project.org'
))
install.packages("mlr3proba")
```

or

```{r eval=FALSE}
install.packages("mlr3proba", repos = "https://mlr-org.r-universe.dev")
```

### Or for easier installation going forward:

1. Run `usethis::edit_r_environ()` then in the file that opened add or edit `options` to look something like

```{r eval=FALSE}
options(repos = c(
       raphaels1 = "https://raphaels1.r-universe.dev",
       mlrorg = "https://mlr-org.r-universe.dev",
       CRAN = 'https://cloud.r-project.org'
))
```

2. Save and close the file, restart your R session
3. Run `install.packages("mlr3proba")` as usual

### Install from GitHub:

```{r eval=FALSE}
remotes::install_github("mlr-org/mlr3proba")
```

## Learners

Core learners are implemented in [mlr3proba](https://mlr3proba.mlr-org.com/reference/index.html#survival-learners), recommended common learners are implemented in [mlr3learners](https://github.com/mlr-org/mlr3learners), and many more are implemented in [mlr3extralearners](https://github.com/mlr-org/mlr3extralearners). Use the [interactive search table](https://mlr-org.com/learners.html) to search for available survival learners and see the [learner status page](https://mlr3extralearners.mlr-org.com/articles/learner_status.html) for their live status.

## Measures

For density estimation only the log-loss is currently implemented, for survival analysis, see full list [here](https://mlr3proba.mlr-org.com/reference/index.html#survival-measures).
Some commonly used measures are the following:

| ID | Measure | Package | Type |
| :--| :------ | :------ | :------ |
| [surv.dcalib](https://mlr3proba.mlr-org.com/reference/mlr_measures_surv.dcalib.html) | D-Calibration | [mlr3proba](https://CRAN.R-project.org/package=mlr3proba) | Calibration
| [surv.cindex](https://mlr3proba.mlr-org.com/reference/mlr_measures_surv.cindex.html) | Concordance Index | [mlr3proba](https://CRAN.R-project.org/package=mlr3proba) | Discrimination
| [surv.uno_auc](https://mlr3proba.mlr-org.com/reference/mlr_measures_surv.uno_auc.html) | Uno's AUC | [survAUC](https://CRAN.R-project.org/package=survAUC) | Discrimination
| [surv.graf](https://mlr3proba.mlr-org.com/reference/mlr_measures_surv.graf.html) | Integrated Brier Score | [mlr3proba](https://CRAN.R-project.org/package=mlr3proba) | Scoring Rule
| [surv.rcll](https://mlr3proba.mlr-org.com/reference/mlr_measures_surv.rcll.html) | Right-Censored Log loss | [mlr3proba](https://CRAN.R-project.org/package=mlr3proba) | Scoring Rule
| [surv.intlogloss](https://mlr3proba.mlr-org.com/reference/mlr_measures_surv.intlogloss.html) | Integrated Log Loss | [mlr3proba](https://CRAN.R-project.org/package=mlr3proba) | Scoring Rule

## Bugs, Questions, Feedback

**mlr3proba** is a free and open source software project that
encourages participation and feedback. If you have any issues,
questions, suggestions or feedback, please do not hesitate to open an
"issue" about it on the [GitHub page](https://github.com/mlr-org/mlr3proba/issues)\!

  In case of problems / bugs, it is often helpful if you provide a
"minimum working example" that showcases the behavior (but don't worry about this if the bug is obvious).


## Similar Projects

Predecessors to this package are previous instances of survival modelling in **[mlr](https://github.com/mlr-org/mlr)**.
The **[skpro](https://github.com/alan-turing-institute/skpro)** package in the python/scikit-learn ecosystem follows a similar interface for probabilistic supervised learning and is an architectural predecessor.
Several packages exist which allow probabilistic predictive modelling with a Bayesian model specific general interface, such as **[rjags](https://CRAN.R-project.org/package=rjags)** and **[stan](https://github.com/stan-dev/rstan)**.
For implementation of a few survival models and measures, a central package is **[survival](https://github.com/therneau/survival)**.
There does not appear to be a package that provides an architectural framework for distribution/density estimation, see **[this list](https://vita.had.co.nz/papers/density-estimation.pdf)** for a review of density estimation packages in R.

## Acknowledgements

Several people contributed to the building of `mlr3proba`. Firstly, thanks to Michel Lang for writing `mlr3survival`. Several learners and measures implemented in `mlr3proba`, as well as the prediction, task, and measure surv objects, were written initially in `mlr3survival` before being absorbed into `mlr3proba`. Secondly thanks to Franz Kiraly for major contributions towards the design of the proba-specific parts of the package, including compositors and predict types. Also for mathematical contributions towards the scoring rules implemented in the package. Finally thanks to Bernd Bischl and the rest of the mlr core team for building `mlr3` and for many conversations about the design of `mlr3proba`.

## Citing mlr3proba

If you use mlr3proba, please cite our [Bioinformatics article](https://doi.org/10.1093/bioinformatics/btab039):
```{r echo = FALSE, comment = ""}
toBibtex(citation("mlr3proba"))
```
